---
title: "fp - EDA"
author: "Ismat"
date: "2024-12-04"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(gridExtra)
library(GGally)
library(pheatmap)
library(ggplot2)
library(patchwork)
library(dplyr)
library(knitr)
library(kableExtra)
library(glmnet)
library(tidyselect)
library(caret)
```

# Data 

```{r}
# Load the dataset
h_data <- read.csv("C:/Users/Ismat/OneDrive/Desktop/fall24/Stat consulting/Final Project/2021_USA_HOSPITAL_DATA.csv")
colnames(h_data)
str(h_data)
```

# EDA techniques to explore the relationship between the categorical variables

```{r,warning=FALSE}
# Change the class of the variable from integer to character
h_data$overall_hospital_rating_star_rating <- as.character(h_data$overall_hospital_rating_star_rating)


plot1 = ggplot(h_data, aes(x = teaching_status, y = malpractice_premiums_per_inpatient_day)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Mal Prem by Teaching Status", x = "Teaching Status", y = "Malpractice Premiums")

plot2 = ggplot(h_data, aes(x = urban_rural, y = malpractice_premiums_per_inpatient_day)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Mal Prem by Urban/Rural", x = "Urban/Rural", y = "Malpractice Premiums")

plot3 = ggplot(h_data, aes(x = bed_size_simple, y = malpractice_premiums_per_inpatient_day)) +
  geom_boxplot(fill = "lightyellow") +
  labs(title = "Mal Prem by bed size", x = "Bed size", y = "Malpractice Premiums")

plot4 = ggplot(h_data, aes(x = ownership_type_simple, y = malpractice_premiums_per_inpatient_day)) +
  geom_boxplot(fill = "yellow") +
  labs(title = "Mal Prem by ownership type", x = "Ownership type", y = "Malpractice Premiums")

plot5 = ggplot(h_data, aes(x = overall_hospital_rating_star_rating, y = malpractice_premiums_per_inpatient_day)) +
  geom_boxplot(fill = "green") +
  labs(title = "Mal Prem by hospital rating", x = "hospital rating", y = "Malpractice Premiums")
grid.arrange(plot1, plot2, plot3, plot4, plot5, ncol = 3, nrow = 2)


```

```{r}
# Change the class of the variable from integer to character
h_data$overall_hospital_rating_star_rating <- as.character(h_data$overall_hospital_rating_star_rating)

p1 = h_data %>%
  group_by(bed_size_simple) %>%
  summarise(mean_premiums = mean(malpractice_premiums_per_inpatient_day, na.rm = TRUE)) %>%
  ggplot(aes(x = bed_size_simple, y = mean_premiums)) +
  geom_bar(stat = "identity", fill = "coral") +
  labs(title = "Malpractice by Bed Size", x = "Bed Size", y = "Mean Malpractice Premiums")
p2 = h_data %>%
  group_by(overall_hospital_rating_star_rating) %>%
  summarise(mean_premiums = mean(malpractice_premiums_per_inpatient_day, na.rm = TRUE)) %>%
  ggplot(aes(x = overall_hospital_rating_star_rating, y = mean_premiums)) +
  geom_bar(stat = "identity", fill = "lightyellow") +
  labs(title = "Malpractice by star rating", x = "star rating", y = "Mean Malpractice Premiums")
p3 = h_data %>%
  group_by(ownership_type_simple) %>%
  summarise(mean_premiums = mean(malpractice_premiums_per_inpatient_day, na.rm = TRUE)) %>%
  ggplot(aes(x = ownership_type_simple, y = mean_premiums)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Malpractice by ownership type", x = "ownership type", y = "Mean Malpractice Premiums")
p4 =  h_data %>%
  group_by(teaching_status) %>%
  summarise(mean_premiums = mean(malpractice_premiums_per_inpatient_day, na.rm = TRUE)) %>%
  ggplot(aes(x = teaching_status, y = mean_premiums)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Malpractice by teaching", x = "teaching Status", y = "Mean Malpractice Premiums")
p5 =  h_data %>%
  group_by(urban_rural) %>%
  summarise(mean_premiums = mean(malpractice_premiums_per_inpatient_day, na.rm = TRUE)) %>%
  ggplot(aes(x = urban_rural, y = mean_premiums)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Malpractice by urban-rural", x = "urban-rural", y = "Mean Malpractice Premiums")
grid.arrange(p1, p2, p3, p4, p5, ncol = 3, nrow = 2)

```

```{r,warning=FALSE}
# Ensure required libraries are loaded
library(dplyr)
library(ggplot2)
library(tidyr)

# Reshape data to long format using pivot_longer
h_data_long <- h_data %>%
  pivot_longer(cols = c("bed_size_simple", "overall_hospital_rating_star_rating", 
                        "ownership_type_simple", "teaching_status", "urban_rural"),
               names_to = "Variable",
               values_to = "Category")

# Create a bar plot with facet_wrap
facet_plot <- ggplot(h_data_long, aes(x = Category, y = malpractice_premiums_per_inpatient_day)) +
  geom_bar(stat = "summary", fun = "mean", fill = "lightblue") +  # Use mean for each category
  facet_wrap(~ Variable, scales = "free_x") +                      # Facet by the categorical variable
  labs(title = "Malpractice Premiums by Different Categories",
       x = "Category",
       y = "Mean Malpractice Premiums") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

# Display the plot
print(facet_plot)
```

#  correlation between the response and all potential variables

```{r}

# correlation between the response and all potential variables
vars_with_potential <- h_data %>%
  select(malpractice_premiums_per_inpatient_day,cost_of_care_per_inpatient_day,charity_care_burden,bad_debt_burden,salary_cost_per_inpatient_day,alos,occupancy_rate,private_room_percentage,RN_percent,LPN_percent,naor_percent,ma_percent,nursing_hours_per_inpatient_day,special_care_day_rate,SDI_score,veteran,medicaid_caseload,disproportionate_percentage,
cms_caseload,percent_of_service_area_under_public,
percent_of_service_area_under_private,hospital_wide_readmission_rate,heart_failure_readmission_rate,acute_myocardial_infarction_readmission_rate,copd_readmission_rate,hip_knee_replacement_readmission_rate,pneumonia_readmission_rate,clabsi_rate,cauti_rate,ssi_colon_rate,
ssi_hyster_rate,mrsa_rate,cdi_rate,PSI_03,PSI_04,PSI_06,PSI_08,PSI_09,PSI_10,PSI_11,PSI_12)
# Compute the correlation matrix
cor_matrix <- cor(vars_with_potential, use = "complete.obs")


# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = ,cellwidth = 15,cellheight = 8)


```

# To explore the correlation between all readmission variable

```{r}
# Select only readmission variables
continuous_read_vars <- h_data %>%
  select(malpractice_premiums_per_inpatient_day,hospital_wide_readmission_rate,
         heart_failure_readmission_rate, acute_myocardial_infarction_readmission_rate,
         copd_readmission_rate, hip_knee_replacement_readmission_rate,
         pneumonia_readmission_rate)

# Compute the correlation matrix
cor_matrix <- cor(continuous_read_vars, use = "complete.obs")

# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = 10,cellwidth = 25,cellheight = 8)

```
# correlation between the variables with PSI
```{r}

# correlation between the variables with PSI
vars_with_psi <- h_data %>%
  select(clabsi_rate,cauti_rate,ssi_colon_rate,
ssi_hyster_rate,mrsa_rate,cdi_rate,
PSI_03,PSI_04,PSI_06,PSI_08,PSI_09,PSI_10,PSI_11,
PSI_12,PSI_13,PSI_14,PSI_15,PSI_90,malpractice_premiums_per_inpatient_day)
# Compute the correlation matrix
cor_matrix <- cor(vars_with_psi, use = "complete.obs")



# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = 7,cellwidth = 20,cellheight = 5)


```

```{r}
# correlation between the variables with sdi score related variables
vars_with_other <- h_data %>%
  select(malpractice_premiums_per_inpatient_day,
SDI_score,
at_least_high_school,
under_100_FPL,
renting,
crowded,
no_computer_internet_access,
not_proficient_english_adult,
unemployment,
non_employment,
single_parent_family,
no_vehicle_households,
long_commute,
veteran,
medicaid_caseload,
disproportionate_percentage,
cms_caseload,
percent_of_service_area_under_public,
percent_of_service_area_under_private,

)
# Compute the correlation matrix
cor_matrix <- cor(vars_with_other, use = "complete.obs")


# Save the plot as a PNG file
#png("correlation_heatmap.png", width = 800, height = 600)  # Specify dimensions and filename

# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = 7,cellwidth = 16,cellheight = 8)

# Close the device to save the file
#dev.off()
```

```{r}
# correlation between the variables with sdi score related variables
vars_with_other <- h_data %>%
  select(malpractice_premiums_per_inpatient_day,
charity_care_burden,
bad_debt_burden,
salary_cost_per_inpatient_day,
malpractice_premiums_per_inpatient_day,
med_records_library_per_inpatient_day,
mean_drg_weighting,
alos,
occupancy_rate,
private_room_percentage,
semi_private_room_percentage,
RN_percent,
LPN_percent,
naor_percent,
ma_percent,
nursing_hours_per_inpatient_day,
special_care_day_rate

)
# Compute the correlation matrix
cor_matrix <- cor(vars_with_other, use = "complete.obs")
                  
# Save the plot as a PNG file
#png("correlation_heatmap.png", width = 800, height = 600)  # Specify dimensions and filename

# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = 7,cellwidth = 16,cellheight = 8)

# Close the device to save the file
#dev.off()

```

```{r}
# correlation between the variables with sdi score related variables
vars_with_cost <- h_data %>%
  select(malpractice_premiums_per_inpatient_day,
charity_care_burden,
bad_debt_burden,
salary_cost_per_inpatient_day,
occupancy_rate,
special_care_day_rate,
med_records_library_per_inpatient_day
)
# Compute the correlation matrix
cor_matrix <- cor(vars_with_cost, use = "complete.obs")
                  
# Save the plot as a PNG file
#png("correlation_heatmap.png", width = 800, height = 600)  # Specify dimensions and filename

# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = 7,cellwidth = 16,cellheight = 8)

# Close the device to save the file
#dev.off()

```
# (IMp)Last correlation matrix with most significant variables after analyzing other correlation matrix

```{r}
# correlation between the response and most significant variables 
vars_significant <- h_data %>%
  select(SDI_score,charity_care_burden,salary_cost_per_inpatient_day,cost_of_care_per_inpatient_day,occupancy_rate,malpractice_premiums_per_inpatient_day,alos,med_records_library_per_inpatient_day,RN_percent,naor_percent,ma_percent,nursing_hours_per_inpatient_day,veteran,long_commute,no_vehicle_households,medicaid_caseload,
disproportionate_percentage,cms_caseload,percent_of_service_area_under_public,percent_of_service_area_under_private,PSI_08, PSI_09, PSI_10, PSI_11, PSI_12 )
# Compute the correlation matrix
cor_matrix <- cor(vars_significant, use = "complete.obs")

# Save the plot as a PNG file
#png("correlation_heatmap.png", width = 800, height = 600)  # Specify dimensions and filename

# Plot the heatmap
pheatmap(cor_matrix, 
         display_numbers = TRUE,  # Show correlation values in cells
         color = colorRampPalette(c("red", "white", "blue"))(100),  # Color scale
         main = "Correlation Heatmap of Continuous Variables with Response Variable",fontsize = 7,cellwidth = 16,cellheight = 8)

# Close the device to save the file
#dev.off()
```




# response analysis

```{r}
summary(h_data$malpractice_premiums_per_inpatient_day)
#Distribution Plot: Plot a histogram or density plot to visualize the distribution of malpractice_premiums_per_inpatient_day. This will help in identifying whether the distribution is skewed, normal, or has outliers.

ggplot(h_data, aes(x = malpractice_premiums_per_inpatient_day)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Malpractice Premiums per Inpatient Day")

ggplot(h_data, aes(y = malpractice_premiums_per_inpatient_day)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Boxplot of Malpractice Premiums per Inpatient Day",
       y = "Malpractice Premiums per Inpatient Day") +
  theme_minimal()

# Save the plot
ggsave("malpractice_boxplot.png", width = 8, height = 6)


```




```{r,warning=FALSE}
# Reshape data to long format using tidyr
library(ggplot2)
library(tidyr)

# Gather predictors into a long format
h_data_long <- h_data %>%
  pivot_longer(cols = c("cost_of_care_per_inpatient_day", "salary_cost_per_inpatient_day", 
                        "med_records_library_per_inpatient_day", "special_care_day_rate", 
                        "charity_care_burden", "bad_debt_burden"),
               names_to = "Predictor",
               values_to = "Value")

# Create scatter plot with facet_wrap
facet_plot <- ggplot(h_data_long, aes(x = Value, y = malpractice_premiums_per_inpatient_day)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +  # Add trendline
  facet_wrap(~ Predictor, scales = "free_x") +            # Facet by predictor
  labs(title = "Relationships Between Predictors and Malpractice Premiums",
       x = "Predictor Value",
       y = "Malpractice Premiums per Inpatient Day") +
  theme_minimal()

# Display the plot
print(facet_plot)

```



# Descriptive statistics
```{r, warning=FALSE}
# Potential variables
selected_vars <- h_data %>%
  select(malpractice_premiums_per_inpatient_day,cost_of_care_per_inpatient_day,charity_care_burden,bad_debt_burden,salary_cost_per_inpatient_day,alos,occupancy_rate,private_room_percentage,RN_percent,LPN_percent,naor_percent,ma_percent,nursing_hours_per_inpatient_day,special_care_day_rate,SDI_score,veteran,medicaid_caseload,disproportionate_percentage,
cms_caseload,percent_of_service_area_under_public,
percent_of_service_area_under_private,hospital_wide_readmission_rate,heart_failure_readmission_rate,acute_myocardial_infarction_readmission_rate,copd_readmission_rate,hip_knee_replacement_readmission_rate,pneumonia_readmission_rate,clabsi_rate,cauti_rate,ssi_colon_rate,
ssi_hyster_rate,mrsa_rate,cdi_rate,PSI_03,PSI_04,PSI_06,PSI_08,PSI_09,PSI_10,PSI_11,PSI_12)

# Calculate descriptive statistics
descriptive_stats <- selected_vars %>%
  summarise(
    Min = sapply(selected_vars, min, na.rm = TRUE),
    Q1 = sapply(selected_vars, quantile, 0.25, na.rm = TRUE),
    Median = sapply(selected_vars, median, na.rm = TRUE),
    Mean = sapply(selected_vars, mean, na.rm = TRUE),
    Q3 = sapply(selected_vars, quantile, 0.75, na.rm = TRUE),
    Max = sapply(selected_vars, max, na.rm = TRUE),
    SD = sapply(selected_vars, sd, na.rm = TRUE)
  ) %>%
  
  as.data.frame() %>%
  `rownames<-`(c( "malpractice_premiums_per_inpatient_day","cost_of_care_per_inpatient_day","charity_care_burden","bad_debt_burden","salary_cost_per_inpatient_day","alos","occupancy_rate","private_room_percentage","RN_percent","LPN_percent","naor_percent","ma_percent","nursing_hours_per_inpatient_day","special_care_day_rate","SDI_score","veteran","Medicaid_Caseload","disproportionate_percentage","cms_caseload","percent_of_service_area_under_public","percent_of_service_area_under_private","hospital_wide_readmission_rate","heart_failure_readmission_rate","acute_myocardial_infarction_readmission_rate","copd_readmission_rate","hip_knee_replacement_readmission_rate","pneumonia_readmission_rate","clabsi_rate","cauti_rate","ssi_colon_rate","ssi_hyster_rate","mrsa_rate","cdi_rate","PSI_03","PSI_04","PSI_06","PSI_08","PSI_09","PSI_10","PSI_11","PSI_12"))

# Format and display the table
descriptive_stats %>%
  kbl(caption = "Descriptive Statistics for Selected Variables", digits = 2) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
# Save as LaTeX table
descriptive_stats %>%
  kbl(format = "latex", caption = "Descriptive Statistics for Selected Variables", digits = 2) %>%
  save_kable("descriptive_stats.tex")
# Save descriptive statistics as a CSV file
#write.csv(descriptive_stats, "descriptive_stats.csv", row.names = TRUE)


```




# Lasso model

```{r}

# Select numeric columns
data_numeric <- h_data %>%
  select_if(is.numeric)

# Remove rows with missing values
data_numeric <- na.omit(data_numeric)

# Prepare the data for Lasso regression
# Set 'malpractice_premiums_per_inpatient_day' as the response variable (y) and the rest as predictors (X)
y <- data_numeric$malpractice_premiums_per_inpatient_day
X <- data_numeric %>% select(-malpractice_premiums_per_inpatient_day)

# Fit the Lasso model
lasso_model <- cv.glmnet(as.matrix(X), y, alpha = 1)

# Get the coefficients from the Lasso model
lasso_coefficients <- coef(lasso_model, s = "lambda.min")

# Print the coefficients
print(lasso_coefficients)

# Identify which variables are selected by Lasso (non-zero coefficients)
selected_variables <- rownames(lasso_coefficients)[lasso_coefficients[,1] != 0]
selected_variables
```




# missing values

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)

# Calculate missing values for each variable
missing_values <- h_data %>%
  summarise_all(~sum(is.na(.))) %>%
  pivot_longer(cols = everything(), 
               names_to = "Variable", 
               values_to = "Missing Values")

# View the missing values table
missing_values

```

# 1. Prepare the Data:

```{r}
# Clean the Data
# List of variables selected by the Lasso model
selected_variables <- c("bad_debt_burden" ,"salary_cost_per_inpatient_day" ,"alos", "private_room_percentage", "semi_private_room_percentage",  "LPN_percent" ,"ma_percent","crowded", "unemployment","no_vehicle_households","long_commute","medicaid_caseload", "percent_of_service_area_under_public", "mrsa_rate","PSI_08" ,"PSI_09",  "PSI_10", "PSI_12", "hospital_wide_readmission_rate","pneumonia_readmission_rate","malpractice_premiums_per_inpatient_day")        

# Select only the relevant columns from your data
data_selected <- h_data %>%select(all_of(selected_variables))


# Check for missing values
sum(is.na(h_data))

# Impute missing values
library(mice)
set.seed(123)  # For reproducibility
imputed_data <- mice(data_selected, m = 1, method = 'pmm', maxit = 5)
data_cleaned <- complete(imputed_data)

# Verify that there are no missing values
sum(is.na(data_cleaned))


```

# Data preparation for modelling

```{r}
#  Split the Data


# Set seed for reproducibility
set.seed(123)

# Split data into training (70%) and testing (30%) sets
trainIndex <- createDataPartition(data_cleaned$malpractice_premiums_per_inpatient_day, p = 0.8, list = FALSE)
train_data <- data_cleaned[trainIndex, ]
test_data <- data_cleaned[-trainIndex, ]
dim(train_data)
dim(test_data)
```

# Decision tree model

```{r}
# Load necessary library
library(rpart)
library(rpart.plot)

# Train a Decision Tree
decision_tree_model <- rpart(malpractice_premiums_per_inpatient_day ~ ., data = train_data, method = "anova")

# Plot the tree
rpart.plot(decision_tree_model)

# Predict and evaluate
predictions_tree <- predict(decision_tree_model, newdata = test_data)
RMSE_tree <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_tree)^2))
print(paste("Decision Tree RMSE:", RMSE_tree))

```

```{r decision tree}
# Load required libraries
library(tidymodels)
library(partykit)

# Fit the decision tree model
decision_tree_model <- decision_tree(mode = "regression") |>
  set_engine("rpart") |>
  fit(malpractice_premiums_per_inpatient_day ~ ., data = train_data)  # Fixed the formula syntax

# Print the decision tree model
print(decision_tree_model)

# Plot the decision tree
plot(as.party(decision_tree_model$fit))

# Predict and evaluate
predictions_tree <- predict(decision_tree_model, new_data = test_data)
RMSE_tree <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_tree$.pred)^2))  # Adjust prediction column

# Print RMSE
print(paste("Decision Tree RMSE:", RMSE_tree))



```

# Random Forest Model

```{r}
# Load necessary library
library(randomForest)

# Train a Random Forest
set.seed(123)
rf_model <- randomForest(malpractice_premiums_per_inpatient_day ~ ., data = train_data, importance = TRUE)

# Plot feature importance
varImpPlot(rf_model)

# Predict and evaluate
predictions_rf <- predict(rf_model, newdata = test_data)
RMSE_rf <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_rf)^2))
print(paste("Random Forest RMSE:", RMSE_rf))
importance(rf_model)

```


# Neural network model

```{r}
# Load necessary library
library(neuralnet)

# Normalize data
normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}
data_normalized <- as.data.frame(lapply(data_cleaned, normalize))

# Split normalized data
train_data_norm <- data_normalized[trainIndex, ]
test_data_norm <- data_normalized[-trainIndex, ]

# Train Neural Network
nn_model <- neuralnet(malpractice_premiums_per_inpatient_day ~ ., data = train_data_norm, hidden = c(5, 3), linear.output = TRUE)

# Plot the neural network
plot(nn_model)

# Predict and evaluate
predictions_nn <- compute(nn_model, test_data_norm[, -1])$net.result
predictions_nn <- predictions_nn * (max(data_cleaned$malpractice_premiums_per_inpatient_day) - min(data_cleaned$malpractice_premiums_per_inpatient_day)) + min(data_cleaned$malpractice_premiums_per_inpatient_day)
RMSE_nn <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_nn)^2))
print(paste("Neural Network RMSE:", RMSE_nn))

```


# Compare model

```{r}
print(paste("Decision Tree RMSE:", RMSE_tree))
print(paste("Random Forest RMSE:", RMSE_rf))
print(paste("Neural Network RMSE:", RMSE_nn))

```


#  Hyperparameter Tuning

## Decision Tree


```{r}
# Load the caret package
library(caret)

# Set up cross-validation
set.seed(123)
control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# Train a Decision Tree with hyperparameter tuning
decision_tree_tuned <- train(
  malpractice_premiums_per_inpatient_day ~ .,
  data = train_data,
  method = "rpart",
  trControl = control,
  tuneLength = 10  # Explore 10 different hyperparameter combinations
)

# Best model and parameters
print(decision_tree_tuned$bestTune)

# Plot the results
plot(decision_tree_tuned)

# Evaluate the model on the test data
predictions_tree_tuned <- predict(decision_tree_tuned, newdata = test_data)
RMSE_tree_tuned <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_tree_tuned)^2))
print(paste("Tuned Decision Tree RMSE:", RMSE_tree_tuned))

```

## Random Forest

```{r}
# Train a Random Forest with hyperparameter tuning
set.seed(123)
rf_tuned <- train(
  malpractice_premiums_per_inpatient_day ~ .,
  data = train_data,
  method = "rf",
  trControl = control,
  tuneGrid = expand.grid(mtry = seq(2,  ncol(train_data) - 1, by = 2))  # Tune the mtry parameter
)

# Best model and parameters
print(rf_tuned$bestTune)

# Plot the results
plot(rf_tuned)
# Plot feature importance
varImpPlot(rf_tuned$finalModel, main = "Feature Importance (Random Forest)")

# Evaluate the model on the test data
predictions_rf_tuned <- predict(rf_tuned, newdata = test_data)
RMSE_rf_tuned <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_rf_tuned)^2))
print(paste("Tuned Random Forest RMSE:", RMSE_rf_tuned))

```

## Neural Network


```{r}
# Train a Neural Network with hyperparameter tuning
set.seed(123)
nn_tuned <- train(
  malpractice_premiums_per_inpatient_day ~ .,
  data = train_data_norm,
  method = "nnet",
  trControl = control,
  tuneGrid = expand.grid(size = c(3, 5, 7), decay = c(0.1, 0.01, 0.001)),  # Tune hidden layers and regularization
  linout = TRUE
)

# Best model and parameters
print(nn_tuned$bestTune)

# Plot the results
plot(nn_tuned)

# Evaluate the model on the test data
predictions_nn_tuned <- predict(nn_tuned, newdata = test_data_norm)
RMSE_nn_tuned <- sqrt(mean((test_data$malpractice_premiums_per_inpatient_day - predictions_nn_tuned)^2))
print(paste("Tuned Neural Network RMSE:", RMSE_nn_tuned))

```



# Compare Models


```{r}
# Create a summary of RMSEs
rmse_summary <- data.frame(
  Model = c("Decision Tree", "Random Forest", "Neural Network"),
  RMSE = c(RMSE_tree_tuned, RMSE_rf_tuned, RMSE_nn_tuned)
)

print(rmse_summary)

```





